Progress Report on the ``Adult Database'' analysis

### 2009-11-11 report/paper.tex
report translated to english.

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{url}
\usepackage{longtable}
\usepackage[portuguese]{babel}
\usepackage[latin1]{inputenc}
\usepackage{subfigure}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{Descrição do dataset ``Adult Database''}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL
    \mails}
\maketitle

\begin{abstract}
Usado de forma extremamente frequente em testes experimentais relacionados com
data mining, a ``Adult Database'' reveste-se de algumas propriedades
interessantes que a tornam bastante útil para o teste da eficácia de algoritmos
de classificação. Este relatório pretende descrever a ``Adult Database'' de um
ponto de vista estatístico e tendo em conta o seu suporte para tarefas de data
mining.
\end{abstract}

\section{Introdução}

\section{Formato dos Dados}

O ``Adult Database'' consiste de 48842 entradas de informação de pessoas
recolhida durante um censos de 1994 realizado nos Estados Unidos da América. Dos
dados globais, foi extraída uma amostra representativa de pessoas com idades
entre os 16 e os 100 anos e um peso superior a 1.

Cada entrada é composta por um conjunto de 14 atributos e por uma classe,
identificando-a como uma pessoa que ganha mais ou menos de 50K. Os atributos e
respectivos tipos são:

\begin{itemize}
  \item{\textbf{Age}: idade do indíviduo, sendo um tipo de dados contínuo.}
  \item{\textbf{Workclass}: Classe operária do indivíduo, sendo um tipo de dados
        categórico: [[Private], [Self-emp-not-inc], [Self-emp-inc], 
        [Federal-gov], [Local-gov], [State-gov], [Without-pay], [Never-worked]]}
  \item{\textbf{FNLWGT}: Medida demográfico do indíviduo. Sendo um tipo de dados
        contínuo, deve ser próximo entre indivíduos com as mesmas
        características sócio-económicas.}
  \item{\textbf{Education}: Nível educacional do indivíduo, sendo um tipo de
        dados ordinal: [1. [Preschool], 2. [1st-4th], 3. [5th-6th], 
        4. [7th-8th], 5. [9th < 10th], 6. [11th < 12th], 7. [HS-grad], 
        8. [Prof-school], 9. [Assoc-acdm], 10. [Assoc-voc], 11. [Some-college],
        12. [Bachelors], 13. [Masters], 14. [Doctorate]]}
  \item{\textbf{Education-num}: Representação numérica do atributo
        \textbf{Education}, sendo um tipo de dados contínuo.}
  \item{\textbf{Marital-status}: Estado civil do indivíduo, sendo um tipo de
        dados categórico: [[Married-civ-spouse], [Divorced], [Never-married],
        [Separated], [Widowed], [Married-spouse-absent], [Married-AF-spouse]]}
  \item{\textbf{Occupation}: Área de ocupação profissional do indivíduo, sendo
        um tipo de dados categórico: [[Tech-support], [Craft-repair],
        [Other-service], [Sales], [Exec-managerial], [Prof-specialty], 
        [Handlers-cleaners], [Machine-op-inspct], [Adm-clerical], 
        [Farming-fishing], [Transport-moving], [Priv-house-serv], 
        [Protective-serv], [Armed-Forces]]}
  \item{\textbf{Relationship}: Tipo de relação familiar do indivíduo, sendo um
        tipo de dados categórico: [[Wife], [Own-child], [Husband], 
        [Not-in-family], [Other-relative], [Unmarried]]}
  \item{\textbf{Race}: Raça do indivíduo, sendo um tipo de dados categórico: 
        [[White], [Asian-Pac-Islander], [Amer-Indian-Eskimo], [Other], [Black]]}
  \item{\textbf{Sex}: Sexo do indivíduo, sendo um tipo de dados categórico:
        [[Male], [Female]]}
  \item{\textbf{Capital-gain}: Ganho de capital do indivíduo, sendo um tipo de
        dados contínuo}
  \item{\textbf{Capital-loss}: Perda de capital do indivíduo, sendo um tipo de
        dados contínuo}
  \item{\textbf{Hours-per-week}: Horas de trabalho por semana do indivíduo,
        sendo um tipo de dados contínuo}
  \item{\textbf{Native-country}: Naturalidade do indivíduo, sendo um tipo de
        dados categórico.}
\end{itemize}

\section{Identificação do problema}

\section{Análise dos Dados}

A tabela \ref{tab:summary} apresenta um sumário da distribuição de valores no
dataset estudado. Numa primeira observação da distribuição de valores, pode-se 
constatar que são poucas as classes com \textit{missing vales} nos dados. Por
outro lado, nalguns atributos, alguns valores têm uma frequência bastante
superior, como são exemplo a naturalidade 'Estados Unidos', a classe operária
'Private' e a raça 'White'. Nas horas de trabalho semanais, verifica-se a
existência de bastantes \textit{outliers}, com valores pouco realistas,
nomeadamente próximos das 100 horas ou menores que 10 em indivíduos com
profissões definidas, como pode ser visto na figura \ref{fig:box_hours}.
Variáveis como o ganho e perda capital apresentam poucos valores superiores a 0,
sendo atributos bastante específicos de determinadas classes de indivíduos. O
sexo também apresenta alguma discrepância na amostra, sendo o número de
indivíduos homens mais do dobro das mulheres.

@@

\caption{Sumário da distribuição de valores no dataset}

@@

\begin{figure}
\centering
\includegraphics[width=11cm]{boxplot_hours_per_week.png}
\caption{Boxplot da variável correspondente às horas de trabalho semanais}
\label{fig:box_hours}
\end{figure} %}

\subsection{Regras de associacao}

\section{Trabalho Futuro}

\section{Conclusões}

>>>>

\documentclass[a4paper]{llncs}
\usepackage{graphicx}
\usepackage{url}
\usepackage{longtable}
\usepackage[latin1]{inputenc}
\usepackage{subfigure}
\usepackage{a4wide}
\urldef{\mails}\path|{ei05011,ei05028}@fe.up.pt|
\bibliographystyle{splncs}
\begin{document}
\title{Progress Report on the ``Adult Database'' analysis}
\author{Flávio Cruz \and João Azevedo}
\institute{Faculdade de Engenharia da Universidade do Porto\\
    Rua Dr. Roberto Frias, s/n 4200-465 Porto PORTUGAL
    \mails}
\maketitle

\begin{abstract}
Highly frequent in experimental tests related to data mining, the Adult Database
has some interesting properties which make it quite useful to test the 
efficiency of classification algorithms. This report aims to describe the Adult
Database from a statistical perspective and taking in account its support for
data mining tasks.
\end{abstract}

\section{Introduction}

\section{Data Format}

The ``Adult Database'' consists of 48842 entries of individual information of 
people gathered in a 1994 census on the United States of America. From the
global collection of data, a representative sample was collected from people
with ages between 16 and 100 and a final weight above 1.

Each entry is composed by a set of 14 attributes and a class, identifying
persons who gain more or less than 50K. The attributes and types are:

\begin{itemize}
  \item{\textbf{Age}: continuous.}
  \item{\textbf{Workclass}: categorical: [[Private], [Self-emp-not-inc], 
        [Self-emp-inc], [Federal-gov], [Local-gov], [State-gov], [Without-pay], 
        [Never-worked]]}
  \item{\textbf{FNLWGT}: continuous. Represents the final weight. People with 
        the same demographical characteristics should have the same weight.}
  \item{\textbf{Education}: ordinal: [1. [Preschool], 2. [1st-4th], 3. 
        [5th-6th], 4. [7th-8th], 5. [9th < 10th], 6. [11th < 12th], 7. 
        [HS-grad], 8. [Prof-school], 9. [Assoc-acdm], 10. [Assoc-voc], 11.
        [Some-college], 12. [Bachelors], 13. [Masters], 14. [Doctorate]]}
  \item{\textbf{Education-num}: continuous. Is a continuous representation of 
        the \textbf{Education} attribute.}
  \item{\textbf{Marital-status}: categorical: [[Married-civ-spouse], [Divorced],
        [Never-married], [Separated], [Widowed], [Married-spouse-absent],
        [Married-AF-spouse]]}
  \item{\textbf{Occupation}: categorical: [[Tech-support], [Craft-repair],
        [Other-service], [Sales], [Exec-managerial], [Prof-specialty], 
        [Handlers-cleaners], [Machine-op-inspct], [Adm-clerical], 
        [Farming-fishing], [Transport-moving], [Priv-house-serv], 
        [Protective-serv], [Armed-Forces]]}
  \item{\textbf{Relationship}: categorical: [[Wife], [Own-child], [Husband], 
        [Not-in-family], [Other-relative], [Unmarried]]}
  \item{\textbf{Race}: categorial: [[White], [Asian-Pac-Islander], 
        [Amer-Indian-Eskimo], [Other], [Black]]}
  \item{\textbf{Sex}: categorical: [[Male], [Female]]}
  \item{\textbf{Capital-gain}: continuous.}
  \item{\textbf{Capital-loss}: continuous.}
  \item{\textbf{Hours-per-week}: continuous.}
  \item{\textbf{Native-country}: categorical.}
\end{itemize}

\section{Problem Identification}

\section{Data Analysis}

The table \ref{tab:summary} summarizes the distribution of values on the studied
dataset. From a first observation, one can see that the amount of missing values
is low. On the other side, in some of the attributes, some classes have a
much higher frequency than others, as the 'united-states' native country, the
'private' workclass and the 'white' race. On the hours of work per week
attribute, there is a great amount of outliers, with some unrealistic
values, namely with values close to the 100 hours or less than 10 in individuals
with a defined occupation, as can be seen on figure \ref{fig:box_hours}.
Variables such has the capital gain and capital loss have few values above 0,
being quite specific to certain classes of individuals. The sex is also
unbalanced in the sample data, as the amount of men is more than twice the
amount of women.

It is also important to verify some correlations between variables. To that
matter, we've tested some apparent situations, such as the influence of the age,
hours of work per week and educational level on the amount gained by an
individual. The figures \ref{fig:age_plus50}, \ref{fig:education_plus50} and
\ref{fig:hpw_plus50} show that there is a tendency for older people, people with
a higher educational level and people who work a great amount of hours per week
to gain a larger amount of money.

\subsection{Association Rules}

\section{Future Work}

\section{Conclusions}

\clearpage

\appendix

\section{Tables}

@@

\caption{Summary of the value distribution in the database}

@@

\clearpage

\section{Figures}

\begin{figure}
\centering
\includegraphics[width=10cm]{boxplot_hours_per_week.png}
\caption{Boxplot of the hours per week variable}
\label{fig:box_hours}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=10cm]{plot_plus_50_age.png}
\caption{Boxplot of the individuals' age, for each class of gain}
\label{fig:age_plus50}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=10cm]{plot_plus_50_education_num.png}
\caption{Boxplot of the individuals' education level, for each class of gain}
\label{fig:education_plus50}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=10cm]{plot_plus_50_hpw.png}
\caption{Boxplot of the individuals' hours of work per week, for each class of 
gain}
\label{fig:hpw_plus50}
\end{figure}
